\textbf{{\Large {Recurrent relation for branching coefficients of affine Lie algebras}}}

### 2009-10-26 AffineLieAlgebras/InformalNotes/article.tex
Minor fix
B4 example added to the article

Let $\gamma
_{0} $ be the lowest vector with respect to the natural ordering in $%
\overset{\circ }{\Delta _{\frak{a}}}$ in the lowest grade of $\Phi _{\frak{a}\subset \frak{g}}$. Decomposing the defining relation 
\begin{equation}
  \label{eq:18}
  \prod_{\alpha\in \pi_{\mathfrak{a}}\circ (\Delta^{+}\setminus \Delta_{\bot})}\left(
    1-e^{-\alpha }\right) ^{\mathrm{{mult}\left( \alpha \right) -{mult}}_{\frak{a%
      }}\mathrm{\left( \alpha \right) }}=-s\left( \gamma _{0}\right) e^{-\gamma
    _{0}}-\sum_{\gamma \in \Phi _{\frak{a}\subset \frak{g}}\setminus \gamma
    _{0}}s\left( \gamma \right) e^{-\gamma },  
\end{equation}
in (\ref{eq:17}) we obtain

\begin{equation}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
  \sum_{\omega\in W/W_{\bot}} \epsilon(\omega) dim\left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right) \delta_{\xi,\pi_{\mathfrak{a}}(\omega(\mu+\rho)-\rho)}+
\sum_{\gamma \in
\Gamma _{\frak{a}\subset \frak{g}}}s\left( \gamma +\gamma _{0}\right) k_{\xi
+\gamma }^{\left( \mu \right) }\right)   \label{recurrent relation}
\end{equation}
where the set 
\begin{equation}
\Gamma _{\frak{a}\subset \frak{g}}=\left\{ \xi -\gamma _{0}|\xi \in \Phi _{%
\frak{a}\subset \frak{g}}\right\} \setminus \left\{ 0\right\} .
\label{fan-defined}
\end{equation}

@@

\section{Algorithm}
\label{sec:algorithm}
\begin{enumerate}
\item Construct the set of anomalous points of the algebra $\mathfrak{g}$
\item Calculate the dimensions of the corresponding modules $L_{\mathfrak{a}_{\bot}}$.
\item Calculate anomalous branching coefficients in the main Weyl chamber of the subalgebra $\mathfrak{a}$
\item \dots
\item PROFIT !!!
\end{enumerate}

@@

Let's describe the reduction of fundamental representation of $B_2$ with the highest weight (in fundamental weight basis) equal to $(1,0)$.
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=80mm]{B2_A1.pdf}
  }
  \caption{Regular embedding of $A_1$ into $B_2$}
  \label{fig:B2_A1}
\end{figure}
On the figure \ref{fig:B2_A1} we have also shown the set of points $\omega(\mu+\rho),\; \omega\in W$ of fundamental representation of $B_2$ with the corresponding determinants of Weyl reflections $\epsilon(\omega)$. 
Now we have to factorise the Weyl group $W$ by $W_{\bot}=\left\{\omega_1\right\}$. We get the following set of anomalous points $\omega(\mu+\rho)-\rho,\; \omega\in W/W_{\bot}$:
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=80mm]{B2_A1_2.pdf}
  }
  \caption{Anomalous points and the corresponding modules}
  \label{fig:B2_A1_2}
\end{figure}
We have also depicted the corresponding $\mathfrak{a}_{\bot}=A_1$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
Then we project these points and dimensions of modules onto the root space of subalgebra $\mathfrak{a}=A_1$ and get the following anomalous points in fundamental weights basis with corresponding multiplicities:
\begin{equation}
  \label{eq:1}
  (1,2),\; (0,-3),\; (-4,3),\; (-5,-2)
\end{equation}
For the function $s(\gamma)$ and the set $\Gamma$ from the definition (\ref{phi-d},\ref{fan-defined}) we have
\begin{equation}
  \label{eq:22}
  (0,1),\; (1,2),\; (2,-1)
\end{equation}
Here the second component denotes the value of $s(\gamma)$.

Anomalous branching coefficient $k^{(1,0)}_{1}=2$, then for anomalous branching coefficient $k^{(1,0)}_{0}$ the formula (\ref{recurrent relation}) gives us
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0}_1 - 3\cdot \delta_{0,0} = 1
\end{equation}
So we've computed the branching coefficients.
\subsubsection{Something high-dimensional}
\label{sec:someth-high-dimens}

>>>>

Let $\gamma
_{0} $ be the lowest vector with respect to the natural ordering in $%
\overset{\circ }{\Delta _{\frak{a}}}$ in the lowest grade of $\Phi _{\frak{a}\subset \frak{g}}$. Decomposing the defining relation 
\begin{equation}
  \label{eq:18}
  \prod_{\alpha\in \pi_{\mathfrak{a}}\circ (\Delta^{+}\setminus \Delta_{\bot})}\left(
    1-e^{-\alpha }\right) ^{\mathrm{{mult}\left( \alpha \right) -{mult}}_{\frak{a%
      }}\mathrm{\left( \alpha \right) }}=-s\left( \gamma _{0}\right) e^{-\gamma
    _{0}}-\sum_{\gamma \in \Phi _{\frak{a}\subset \frak{g}}\setminus \left\{\gamma_{0}\right\}}s\left( \gamma \right) e^{-\gamma },  
\end{equation}
in (\ref{eq:17}) we obtain

\begin{equation}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
  \sum_{\omega\in W/W_{\bot}} \epsilon(\omega) dim\left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right) \delta_{\xi-\gamma_0,\pi_{\mathfrak{a}}(\omega(\mu+\rho)-\rho)}+
\sum_{\gamma \in
\Gamma _{\frak{a}\subset \frak{g}}}s\left( \gamma +\gamma _{0}\right) k_{\xi
+\gamma }^{\left( \mu \right) }\right)   \label{recurrent relation}
\end{equation}
where the set 
\begin{equation}
\Gamma _{\frak{a}\subset \frak{g}}=\left\{ \xi -\gamma _{0}|\xi \in \Phi _{%
\frak{a}\subset \frak{g}}\right\} \setminus \left\{ 0\right\} .
\label{fan-defined}
\end{equation}

@@

\section{Algorithm}
\label{sec:algorithm}
\begin{enumerate}
\item Construct the set of anomalous points of the algebra $\mathfrak{g}$
\item Calculate the dimensions of the corresponding modules $L_{\mathfrak{a}_{\bot}}$.
\item Calculate anomalous branching coefficients in the main Weyl
  chamber of the subalgebra $\mathfrak{a}$ using recurrent relation (\ref{recurrent relation}).
\end{enumerate}

@@

Let's describe the reduction of fundamental representation of $B_2$ with the highest weight (in fundamental weight basis) equal to $(1,0)$.
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=90mm]{B2_A1.pdf}
  }
  \caption{Regular embedding of $A_1$ into $B_2$}
  \label{fig:B2_A1}
\end{figure}
On the figure \ref{fig:B2_A1} we have also shown the set of points $\omega(\mu+\rho),\; \omega\in W$ of fundamental representation of $B_2$ with the corresponding determinants of Weyl reflections $\epsilon(\omega)$. 
Now we have to factorise the Weyl group $W$ by $W_{\bot}=\left\{\omega_1\right\}$. We get the following set of anomalous points $\omega(\mu+\rho)-\rho,\; \omega\in W/W_{\bot}$:
\begin{figure}[ph]
  \noindent\centering{
    \includegraphics[width=90mm]{B2_A1_2.pdf}
  }
  \caption{Anomalous points and the corresponding $\mathfrak{a}_{\bot}=A_1$-modules}
  \label{fig:B2_A1_2}
\end{figure}
We have also depicted the corresponding $\mathfrak{a}_{\bot}=A_1$-modules $L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}$.
Then we project these points and dimensions of modules onto the root space of subalgebra $\mathfrak{a}=A_1$ and get the following anomalous points in fundamental weights basis with corresponding multiplicities:
\begin{equation}
  \label{eq:1}
  (1,2),\; (0,-3),\; (-4,3),\; (-5,-2)
\end{equation}
For the function $s(\gamma)$ and the set $\Gamma$ from the definition (\ref{phi-d},\ref{fan-defined}) we have
\begin{equation}
  \label{eq:22}
  (0,1),\; (1,2),\; (2,-1)
\end{equation}
Here the second component denotes the value of $s(\gamma)$.

Anomalous branching coefficient $k^{(1,0)}_{1}=2$, then for anomalous branching coefficient $k^{(1,0)}_{0}$ the formula (\ref{recurrent relation}) gives us
\begin{equation}
  \label{eq:23}
  k^{(1,0)}_{0}=-1\cdot k^{(1,0)}_2 +2\cdot k^{(1,0}_1 - 3\cdot \delta_{0,0} = 1
\end{equation}
So we've computed the branching coefficients.
\subsubsection{Embedding of $B_2$ into $B_4$}
\label{sec:someth-high-dimens}
Consider the the regular embedding of the subalgebra $B_2$ into the algebra $B_4$.
We calculate the branching coefficients for the fundamental representation of $B_4$.
The corresponding Dynkin diagrams are in the figure.
\begin{figure}[ph]
  \centering
  \includegraphics[width=60mm]{B4_B2_2A1.pdf}
  \caption{Dynkin diagrams}
  \label{fig:dynkin}
\end{figure}

In the orthogonal basis $e_1,\dots,e_4$ simple roots of $B_4$ are
\begin{equation}
  \label{eq:1}
  (e_1 - e_2,\; e_2 - e_3,\; e_3 - e_4,\; e_4)
\end{equation}
Positive roots are
\begin{multline}
  \label{eq:2}
  (e_1 - e_2,\; e_2 - e_3,\; e_3 - e_4,\; e_4,\; e_1 - e_3,\; e_2 - e_4,\; e_3 + e_4,\; e_3,\; e_1 - e_4,\;\\
    e_2 + e_4,\; e_2,\; e_1 + e_4,\; e_2 + e_3,\; e_1,\; e_1 + e_3,\; e_1 + e_2)
\end{multline}
Simple roots of the embedded subalgebra $\mathfrak{a}=B_2$ are
\begin{equation}
  \label{eq:3}
  (e_3-e_4,e_4)
\end{equation}

The set $\Delta_{\bot}$ is equal to
\begin{equation}
  \label{eq:4}
  \left\{e_1-e_2,e_1+e_2,e_1,e_2\right\}
\end{equation}

We see that this is the set of positive roots of the algebra $\mathfrak{a}_{\bot}=B_2$.

To find the branching coefficients we need to compute the anomalous points of $B_4$, select point lying in the main Weyl chamber of $\mathfrak{a}_{\bot}$ and compute the dimensions of corresponding $\mathfrak{a}_{\bot}$-modules.

We consider the $B_4$-module with the highest weight $\mu=(0,1,0,2)=2
e_1 + 2 e_2 + e_3 + e_4$.

The set of the anomalous points $\omega(\mu+\rho)-\rho,\; \omega\in W$
consists of 384 points. We don't show it here.

We need to select those points $\omega(\mu+\rho)$  which are projected into the main chamber of the embedded algebra $\mathfrak{a}_{\bot}$.
It means that scalar product of these points with all the roots from $\Delta_{\bot}$ is non-negative.

To compute dimensions of the corresponding
$\mathfrak{a}_{\bot}$-modules we need to project each selected point
onto the root space $\Delta_{\bot}$ and substract
$\rho_{\mathfrak{a}_{\bot}}$, then use Weyl dimension formula.

We show the result of this procedure on the figure \ref{fig:B4B2anom}.
\begin{figure}[ph]
    \includegraphics[width=170mm]{B4_B2_anom_points.pdf}
  \caption{Anomalous points with the dimensions of corresponding $\mathfrak{a}_{\bot}$-modules.}
  \label{fig:B4B2anom}
\end{figure}

Then we should construct ``the fan'' and use the recurrent relation for the computation of anomalous branching coefficients.

Using the definition (\ref{fan-defined}) we get the following set of
the points $\Gamma$ with the corresponding values $s(\gamma+\gamma_0)$
\begin{figure}[ph]
  \includegraphics[width=150mm]{B4_B2_fan.pdf}
  \caption{Fan for $B_2\subset B_4$}
  \label{fig:B4B2Fan}
\end{figure}
We use the recurrent relation (\ref{recurrent relation}) and get
following branching coefficients:
\begin{equation}
  \label{eq:24}
  \pi_{\mathfrak{a}} \left(ch L^{(0,1,0,2)}_{B_4}\right) = 6 \; ch L^{(0,0)}_{B_2}+ 60
  \; ch L_{B_2}^{(0,2)}+ 30 \; ch L_{B_2}^{(1,0)}+ 19 \; ch L_{B_2}^{(2,0)}+
  40 \; ch L_{B_2}^{(1,2)}+ 10 \; ch L_{B_2}^{(2,2)}
\end{equation}
The dimension of the highest-weight $B_4$-module $L^{(0,1,0,2)}_{B_4}$
is equal to 2772. It is easy to see, that right-hand side of the
equation (\ref{eq:24}) gives the same result.
